Computing properties of independent sets by generic programming tensor networks \thanks{\funding{...}}

### 2022-03-17 paper/paper.tex
update

We propose to view the computation of all these properties from an abstract algebra perspective and use a unified framework,
generic tensor-network, to solve these seeming unrelated properties all.
%We map the computation of these properties into generic tensor network contraction with specially designed tensor element algebra.
Tensor network is also known as the sum product network in probabilistic modeling~\cite{Bishop2006} or einsum in linear algebra libraries such as numpy.
It is composed of a set of tensors and a set of labels for specifying the contraction of tensors,
where the tensor element types are typically restricted to standard number types such as real numbers and complex numbers.
We extend this concept to generic tensor network by generalizing the tensor element type to any type being a commutative semiring.
Given a good contraction order, the space required to contract a tensor network is 2 power the tree width of its line graph~\cite{Markov2008}.
\purple{I’d write out $2^{{\rm tw}(G)}$}\blue{I feel it is hard to explain line graph without texts.}
Recent progress in simulating quantum circuits with random tensor networks~\cite{Gray2021, Pan2021, Kalachev2021} enables us to find such a good contraction order for a tensor network with thousands of tensors in a reasonable time.
\purple{Is the scaling of these algorithms known?}\blue{They are not exact algorithms, the time can be any, typically a few minutes, will it be helpful to mention this?}
When using tensor network for solving one of the best sets with the maximum size,
this time complexity is the same as the dynamic programming algorithm~\cite{Courcelle1990, Fomin2013}.
These two algorithms are related but are not equivalent.
Dynamic programming is more general purposed, while a tensor network has a richer algebraic structure, and this structure is crutial for solution space property computation.
\purple{Not clear how first part and second part of sentence are related}\blue{now? or maybe just removing the comparison with dynamic programming?}
%For sparse graphs, the treewidth is usually much smaller than the number of vertices~\cite{Fomin2006}. Our algorithms, on the other hand, are much more versatile than traditional methods and can be used to compute many other properties than just the MIS size.

@@

\purple{I think the following paragraph could be clarified. When you read the definition
of generic programming, it’s not clear what “in the most general setting” means. My
understanding is it just means that the program will work regardless of the input data
type - it would be good to say this explicitly. I also think the way it is written takes a while
to understand what dynamically versus statically typed languages are. The definition,
“When these languages “see” a new input type, the compiler can recompile the generic
program for the new type.” comes rather late in the paragraph. Also, while I know what
you mean, concepts like “inlining immutable elements” and “cache miss rate” would be
lost on someone only with a physics background.}
This definition of generic programming contains two major aspects: a single program works in the most general setting and efficiency.
To understand the first aspect on generality, suppose we want to write a function that raises an element to a power, $f(x, n) := x^n$.
One can easily write a function for standard number types that computes the power of $x$ in $O \left( \log(n) \right)$ steps using the multiply and square trick.
Generic programming does not require $x$ to be a standard number type,
instead it treats $x$ as an element with an associative multiplication operation $\odot$ and a multiplicative identity $\mymathbb{1}$.
In such a way, when the program takes a matrix as an input, it computes the matrix power without extra efforts.
The second aspect is about performance. For dynamically typed languages such as Python,
one can easily write very general codes, but the efficiency is not guaranteed; for example, the speed of computing the matrix multiplication between two numpy arrays with python objects as elements is much slower than statically typed languages such as C++ and Julia~\cite{Bezanson2012}.
C++ uses templates for generic programming while Julia takes advantage of just-in-time compilation and multiple dispatch.
When these languages ``see'' a new input type, the compiler can recompile the generic program for the new type.
A myriad of optimizations can be done during the compilation, such as inlining immutable elements with fixed sizes in an array to decrease the cache miss rate when accessing data.
In Julia, these inlined arrays can even be compiled to GPU devices for faster computation~\cite{Besard2018}.

@@

It is in general not computational efficient to evaluate \Eq{eq:idp} by directly summing up the $2^{|V|}$ products, which scales exponentially as the number of vertices.
\purple{Perhaps explicitly say what this scales as?}\blue{I am not sure explicitly saying ``scales exponentially as the number of vertices'' is helpful, because treewidth often scales linearly as number of vertices, so tensor network also scales exponentially as the number vertices in most cases.}
The standard approach to evaluate a tensor network is to contract two tensors at a time with a certain order utilizing the associativity and commutativity of tensor contraction.
A fully optimized contraction order can reduce the time complexity significantly, while requiring a minimum space complexity $O(2^{\text{tw}(G)})$~\cite{Markov2008} to store intermediate contraction results.
\purple{Perhaps just say time and space
complexity? Order of magnitude doesn't make sense to me in terms of scaling. Also,
what is the time complexity of finding a good contraction order? Is there a known
average-case complexity? This seems important.}\blue{see added texts below.}
The pairwise tensor contraction also makes it possible to utilize basic linear algebra subprograms (BLAS) functions to speed up the computation for certain tensor element types.
In practise, it is difficult to find an optimal contraction order for large tensor networks because finding the tree width itself is NP-hard.
However, it is easy to find a close to optimal contraction order in typically a few minutes with a heuristic contraction order finding algorithm~\cite{Kourtis2019,Kalachev2021}.
For large scale applications,
it is also possible to slice over certain degrees of freedom to reduce the space complexity, i.e.
enumerating certain degrees of freedom in an external loop so that the space complexity of contracting a smaller tensor network inside the loop can be smaller.

@@

\purple{I have not checked this yet but will
tomorrow. Is it really easy/simple enough to do without writing stuff down? Is this the simplest
example possible? The equations look quite formidable at first glance, and it takes a lot of time
to match the hyperedge labels in the equation to the physical graph.}

@@

One can feed the standard computer numbers types into this tensor network and evaluate the independence polynomial by contracting it numerically.

@@

the coefficients? It's not very explicit this procedure works at this point.}
However, we are more interested in finding the coefficients of this polynomial, because this quantity tells us the counting of independent sets at each size.

@@

The use of the word “however” here is confusing to me because it feels like you’re
saying that the previous technique doesn’t give the coefficients, or it’s very hard to
obtain the coefficients.}

@@

say that you're about to describe this semiring, otherwise it seems like you’re going to
use the general definition of the semiring rather than a specific type of semiring}

@@

In \App{sec:finitefield}, we provide an accurate and memory efficient method to find the polynomial by fitting it in a finite field.
Alternatively, one can use the method in \App{sec:fft} to fit this polynomial more time efficiently with complex number at the cost of precision.

@@

how the compiler can optimize on this data type?}

\section{Maximum independent sets and its counting}\label{sec:counting}
\subsection{Tropical algebra for finding the MIS size and counting MISs}
The MIS size $\alpha(G)$, or the independence number is an important graph characteristics.
The method we use to compute this quantity is based on the following observations.
Let $x=\infty$, the independence polynomial in the previous section becomes
\begin{equation}
I(G, \infty) = \lim_{x\rightarrow \infty}\sum_{k=0}^{\alpha(G)}a_k x^k = a_{\alpha(G)} \infty^{\alpha(G)},
\end{equation}
where the lower-order terms vanish. We can thus replace the polynomial type $a = (a_0, a_1, \ldots, a_k)$ with a new type with two fields: the largest exponent $k$ and its coefficient $a_k$.
\purple{I was initially confused
because I thought 0 = (0, ?) where “?” could be anything positive since it's multiplied
by zero. It confused me that ? could equal $-\infty$ since I thought you defined your ring for
k positive, since a polynomial has positive exponents. If you have negative exponents I
thought it is a Laurent polynomial or something. After re-reading I see why this doesn't
matter, though.}
From this, we can define a new algebra as
\begin{equation}
    \eqname{P1}
\begin{split}
    a_x\infty^x \oplus a_y\infty^y &= \begin{cases}
        (a_x + a_y)\infty^{\max(x,y)}, & x = y\\
        a_y\infty^{\max(x,y)}, & x < y\\
        a_x\infty^{\max(x,y)}, & x > y
    \end{cases}, \\
    a_x\infty^x \odot a_y\infty^y &= a_x a_y\infty^{x+y}\\
    \mymathbb{0} &= 0\infty^{-\infty}\\
    \mymathbb{1} &= 1\infty^{0}.
\end{split}%& \text{\rlap{(P1)}}
\label{eq:countingtropical}
\end{equation}
To implement this algebra programmatically, we create a data type with two fields $(x, a_x)$ to store the MIS size and its counting,
and define the above operations and constants correspondingly.
If one is only interested in finding the MIS size, one can drop the counting field.
The algebra of the exponents becomes the max-plus tropical algebra~\cite{Maclagan2015, Moore2011}:
\begin{equation}\eqname{T}
    \begin{split}
        x \oplus y &= \max(x,y)\\
        x \odot y &= x + y\\
        \mymathbb{0} &= -\infty\\
        \mymathbb{1} &= 0.
    \end{split}\label{eq:tropical}
\end{equation}
Algebra \Eq{eq:tropical} and \Eq{eq:countingtropical} are the same as those used in Liu et al.~\cite{Liu2021} to calculate and count spin glass ground states.
For independent set calculations here, the vertex tensor and edge tensor becomes:
\begin{equation}
    W^{\rm T} = \left(\begin{matrix}
        \mymathbb{1} \\
        \infty
    \end{matrix}\right),   
    \qquad \qquad
        B^{\rm T} = \left(\begin{matrix}
        \mymathbb{1}  & \mymathbb{1} \\
        \mymathbb{1} & \mymathbb{0}
    \end{matrix}\right).
\end{equation}

@@

\begin{equation}
\eqname{SPT}
\begin{split}
    s \oplus t &= (\texttt{SUM}, /, s, t)\\
    s \odot t &= (\texttt{PROD}, /, s, t)\\
    \mymathbb{0} &= (\texttt{ZERO}, /, /, /)\\
    \mymathbb{1} &= (\texttt{LEAF}, 0^{\otimes |V|}, /, /).
\end{split}\label{eq:expr}
\end{equation}
\purple{It might be nice to have a picture of an example tree if you have time. However,
it is pretty clear from the definition, it would just help to visualize. Do you ever use the
zero element?}
The vertex tensor and edge tensor become
\begin{equation}
    W^{\rm SPT}((\texttt{LEAF},\boldsymbol{e}_i,/,/)) = \left(\begin{matrix}
        \mymathbb{1} \\
        (\texttt{LEAF},\boldsymbol{e}_i,/,/)
    \end{matrix}\right),   
    \qquad
        B^{\rm SPT} = \left(\begin{matrix}
        \mymathbb{1}  & \mymathbb{1} \\
        \mymathbb{1} & \mymathbb{0}
    \end{matrix}\right).
\end{equation}

This algebra is a commutative semiring because we define the equivalence of two sum-product trees by comparing their expanded (using \Eq{eq:set}) forms, rather than their storages.
Although one is probably unable to enumerate the configurations represented by this tree due to the extremely large configuration space size,
but he can directly sample configurations from this tree.
This is because once we have this sum-product representation of the configuration space,
one can compute the number of configurations in each branch rigorously and efficiently.
\purple{To clarify - is the sum-product tree now the coefficients
on the polynomial? I am a little confused on how we would, for instance, use this to
compute the MISs}
To generate samples, one just descends this sum product tree to left or right sibling with probability decided by the size of each branch while doing the computation specified by the first field.
\purple{How do we compute the size of each
branch?}

@@

Figure~\ref{fig:benchmark}(a) shows the time and space complexity of tensor network contraction for different graph sizes, where the space complexity is the same as the treewidth of the problem graph.
The contraction order is obtained using the local search algorithm in Ref.~\cite{Kalachev2021}.
In practice, slicing technique~\cite{Kalachev2021} is used for graphs with treewidth greater than 27 to fit the computation into a 16GB memory.
One can see that all the computing times in figure~\ref{fig:benchmark}(b), (c) and (d) have a strong correlation with the treewidth\purple{with the number
of vertices, and therefore the treewidth}.
\purple{Does this say anything about how
optimized the contraction order is? If so, it might be worth explicitly saying. Some of
the quantities in panel d look like they might be superexponential in the number of
vertices (tree-width).}
Among these benchmarks, computational tasks with data types \texttt{T (CPU)},
\texttt{$\mathbbm{R}$ (CPU)}, \texttt{$\mathbbm{R}$ (GPU)}, \texttt{$\mathbbm{C}$ (CPU)}, \texttt{$\mathbbm{C}$ (GPU)} and \texttt{T+bounding (CPU)} can utilize fast BLAS functions to evaluate tensor contractions, hence 
are much faster comparing to non-BLAS element types in the same category.
\purple{It might be worth explicitly saying
the data types the first time they're mentioned (for example, I don't know what T is.
Also, it might be worth adding a couple descriptor words on bounding.}
GPU computes much faster than CPU in all cases when the problem scale is large enough such that the actual computing time is comparable or larger than the launching overhead of CUDA kernels.
Most algebras can be computed on a GPU, except those requiring dynamic sized structures, i.e.\ types with algebra defined in \Eq{eq:polynomial}, \Eq{eq:set}, \Eq{eq:ext-tropical} and \Eq{eq:expr}.
In figure~\ref{fig:benchmark}(c), one can see the Fourier transformation based method is the fastest in computing the independence polynomial,
but it may suffer from round-off errors\purple{Can the round-off error thing be seen from Figure c?}. The finite field (GF$(p)$) approach is the only method that does not have round-off errors and can be run on a GPU.
In figure~\ref{fig:benchmark}(d), one can see the technique to bound the enumeration space improves the performance for more than one order of magnitude in enumerating the MISs.
Bounding can also reduce the memory usage significantly, without which the largest computable graph size is only $\sim150$ on a device with 32GB main memory.

@@

A useful number that can be computed using the number of independent sets is the entropy constant for the hardcore lattice gases on these graphs.
\purple{Why is it useful and
what does it tell us about the physical system?}
For the square grid graphs, it is called the \textit{hard square entropy constant} (\href{https://oeis.org/A085850}{OEIS A085850}), which is defined as $\lim_{L\rightarrow \infty} F(L,L)^{1/L^2}$, where $F(L,L)$ is the number of independent sets of a given lattice dimensions $L \times L$. This quantity arises in statistical mechanics of hard-square lattice gases~\cite{Baxter1980, Pearce1988} and is used to understand phase transitions for these systems. This entropy constant is not known to have an exact representation, but it is accurately known in many digits. Similarly, we can define entropy constants for other lattice gases. In \Fig{fig:hardsquare}, we look at how $F(L,L)^{1/\lfloor pL^2\rceil}$ scales as a function of the grid size $L$ for all types of graphs shown in Figure~\ref{fig:lattices}. Our results match the known results for the non-disordered square grid and King's graphs. For disordered square grid and King's graphs with a filling factor $p=0.8$, we randomly sample 1000 graph instances. To our knowledge, the entropy constants for these disordered graphs have not been studied before. Interestingly, the variations due to different random instances are negligible for this entropy quantity.

@@

\begin{figure}[th]
   \centering
\centerline{\begin{tikzpicture}[]
    \scriptsize
    \node[draw,fill=lime!80,fill opacity=1, text opacity=1.0,ellipse,minimum width=2cm, minimum height=1cm,inner sep=0pt] at (0, 2.5) (R) {$\mathbbm{R}$};
    \def\dx{-3};
    \node[draw,fill=teal!50,fill opacity=1, text opacity=1.0,ellipse,minimum width=5cm, minimum height=3cm,inner sep=0pt] at (\dx, 0) (PN) {\hspace{-3.5cm}Polynomial};
    \node[draw,fill=brown!75,fill opacity=1, text opacity=1.0,ellipse,minimum width=3.5cm, minimum height=1.0cm,inner sep=0pt] at (\dx, 0.7) (P1) {\hspace{-1.0cm}Highest order};
    \node[draw,fill=brown!40,fill opacity=1, text opacity=1.0,ellipse,minimum width=3.5cm, minimum height=1.0cm,inner sep=0pt] at (\dx, -0.4) (P2) {\hspace{-0.8cm}2nd highest order};
    \node[draw,fill=brown,fill opacity=1, text opacity=1.0,ellipse,minimum width=0.8cm, minimum height=0.3cm,inner sep=0pt] at (\dx, 1.0) (T) {};
    \node[draw,fill=brown!50,fill opacity=1, text opacity=1.0,ellipse,minimum width=0.8cm, minimum height=0.3cm,inner sep=0pt] at (\dx, -0.1) (T2) {};
    \node at (\dx, -1.2) {$\ldots$};
    \node[above = 1cm] at (T) (textT) {Tropical};
    \node[below = 2cm, left=4cm] (textT2) {Extended Tropical};
    \draw[black,-latex] (textT) -- (T);
    \draw[black,-latex] (textT2) -- (T2);

>>>>

We propose to view the computation of all these properties from an abstract algebra perspective and use a unified framework,
generic tensor-network, to solve these seeming unrelated properties all.
%We map the computation of these properties into generic tensor network contraction with specially designed tensor element algebra.
Tensor network is also known as the sum product network in probabilistic modeling~\cite{Bishop2006} or einsum in linear algebra libraries such as numpy.
It is composed of a set of tensors and a set of labels for specifying the contraction of tensors,
where the tensor element types are typically restricted to standard number types such as real numbers and complex numbers.
We extend this concept to generic tensor network by generalizing the tensor element type to any type being a commutative semiring.
Given a good contraction order, the space required to contract a tensor network is 2 power the treewidth of its line graph~\cite{Markov2008}.
\purple{I’d write out $2^{{\rm tw}(G)}$}\blue{I feel it is hard to explain line graph without texts.}
Recent progress in simulating quantum circuits with random tensor networks~\cite{Gray2021, Pan2021, Kalachev2021} enables us to find such a good contraction order for a tensor network with thousands of tensors in a reasonable time.
\purple{Is the scaling of these algorithms known?}\blue{They are not exact algorithms, the time can be any, typically a few minutes, will it be helpful to mention this?}
When using tensor network for solving one of the best sets with the maximum size,
this time complexity is the same as the dynamic programming algorithm~\cite{Courcelle1990, Fomin2013}.
These two algorithms are related but are not equivalent.
Dynamic programming is more general purposed, while a tensor network has a richer algebraic structure, and this structure is crutial for solution space property computation.
\purple{Not clear how first part and second part of sentence are related}\blue{now? or maybe just removing the comparison with dynamic programming?}
%For sparse graphs, the treewidth is usually much smaller than the number of vertices~\cite{Fomin2006}. Our algorithms, on the other hand, are much more versatile than traditional methods and can be used to compute many other properties than just the MIS size.

@@

\purple{I think the following paragraph could be clarified. When you read the definition
of generic programming, it's not clear what ``in the most general setting'' means. My
understanding is it just means that the program will work regardless of the input data
type - it would be good to say this explicitly. I also think the way it is written takes a while
to understand what dynamically versus statically typed languages are. The definition,
“When these languages “see” a new input type, the compiler can recompile the generic
program for the new type.” comes rather late in the paragraph. Also, while I know what
you mean, concepts like “inlining immutable elements” and “cache miss rate” would be
lost on someone only with a physics background.}\blue{This definition is from the book I cited, I also do not like ``in the most general setting'', maybe we just use our version?}
This definition of generic programming contains two major aspects: a single program works in the most general setting and efficiency.
To understand the first aspect on generality, suppose we want to write a function that raises an element to a power, $f(x, n) := x^n$.
One can easily write a function for standard number types that computes the power of $x$ in $O \left( \log(n) \right)$ steps using the multiply and square trick.
Generic programming does not require $x$ to be a standard number type,
instead it treats $x$ as an element with an associative multiplication operation $\odot$ and a multiplicative identity $\mymathbb{1}$.
In such a way, when the program takes a matrix as an input, it computes the matrix power without extra efforts.
The second aspect is about performance. For dynamically typed languages such as Python,
one can easily write very general codes, but the efficiency is not guaranteed; for example, the speed of computing the matrix multiplication between two numpy arrays with python objects as elements is much slower than statically typed languages such as C++ and Julia~\cite{Bezanson2012}.
C++ uses templates for generic programming while Julia takes advantage of just-in-time compilation and multiple dispatch.
When these languages ``see'' a new input type, the compiler can recompile the generic program for the new type.
A myriad of optimizations can be done during the compilation, such as inlining immutable elements with fixed sizes in an array to decrease the cache miss rate when accessing data.
In Julia, these inlined arrays can even be compiled to GPU devices for faster computation~\cite{Besard2018}.

@@

It is in general not computational efficient to evaluate \Eq{eq:idp} by directly summing up the $2^{|V|}$ products, which scales exponentially as the number of vertices.
\purple{Perhaps explicitly say what this scales as?}\blue{I am not sure explicitly saying ``scales exponentially as the number of vertices'' is helpful, because treewidth often scales linearly as number of vertices, so tensor network also scales exponentially as the number vertices in most cases.}
The standard approach to evaluate a tensor network is to contract two tensors at a time with a certain order utilizing the associativity and commutativity of tensor contraction.
A fully optimized contraction order can reduce the time complexity significantly, while requiring a minimum space complexity $O(2^{\text{tw}(G)})$~\cite{Markov2008} to store intermediate contraction results.
\purple{Perhaps just say time and space
complexity? Order of magnitude doesn't make sense to me in terms of scaling. Also,
what is the time complexity of finding a good contraction order? Is there a known
average-case complexity? This seems important.}\blue{see added texts below.}
The pairwise tensor contraction also makes it possible to utilize basic linear algebra subprograms (BLAS) functions to speed up the computation for certain tensor element types.
In practise, it is difficult to find an optimal contraction order for large tensor networks because finding the treewidth itself is NP-hard.
However, it is easy to find a close to optimal contraction order in typically a few minutes with a heuristic contraction order finding algorithm~\cite{Kourtis2019,Kalachev2021}.
For large scale applications,
it is also possible to slice over certain degrees of freedom to reduce the space complexity, i.e.
enumerating certain degrees of freedom in an external loop so that the space complexity of contracting a smaller tensor network inside the loop can be smaller.

@@

\purple{I have not checked this yet but will
tomorrow. Is it really easy/simple enough to do without writing stuff down? Is this the simplest
example possible? The equations look quite formidable at first glance, and it takes a lot of time
to match the hyperedge labels in the equation to the physical graph.}\blue{sometimes I am afraid the example is too trivial to be useful. The simplest example to show contraction order matters can be a 3-vertex one.}

@@

By assigning a standard computer number to $x$, one can evaluate this independence polynomial for this specific value by contracting this tensor network directly.

@@

the coefficients? It's not very explicit this procedure works at this point.}\blue{how about now?}
However, instead of evaluating this polynomial for a certain value, we are more interested in finding the coefficients of this polynomial, because this quantity tells us the counting of independent sets at each size.

@@

The use of the word “however” here is confusing to me because it feels like you're
saying that the previous technique doesn't give the coefficients, or it's very hard to
obtain the coefficients.}\blue{Yep, we haven't found the coefficients. I tried to modify words a bit, does it look better now?}

@@

say that you're about to describe this semiring, otherwise it seems like you're going to
use the general definition of the semiring rather than a specific type of semiring}\blue{from $1$ to $\mymathbb{1}$ and $0$ to $\mymathbb{0}$ applies for a general semiring, I haven't change this part, let's discuss to make sure I understand your point.}

@@

One may argue that one can first evaluate this polynomial at different $x$ being a standard number type,
and then apply the gaussian elimination to fit the coefficients of this polynomial.
This seemingly more time and space efficient approach suffers from severe issues in practise.
The data ranges of standard integer types are too small to cover many practical using cases,
while the floating point numbers may have round off error that much larger than the value itself, because the number of sets at different sizes may vary tens or even hundreds of orders and this makes the small coefficients can not be retrieved correctly by fitting.
For the purpose of evaluating these coefficients, we refer readers to \App{sec:finitefield}, where we provide an accurate and memory efficient method to find the polynomial by fitting it in a finite field.
Alternatively, one can use the method in \App{sec:fft} to fit this polynomial more time efficiently with complex numbers with controllable round off errors.
For the purpose of discussion in the main text, let us stick to this less efficient polynomial algebra.

@@

how the compiler can optimize on this data type?}\blue{I added some discussion here to address why we stick to polynomial number types,
for short, it is a dynamic type that can be slow, so we do not use it a lot in practise.
I mention it here because it is the base of the following discussion.}

\section{Maximum independent sets and its counting}\label{sec:counting}
\subsection{Tropical algebra for finding the MIS size and counting MISs}
The MIS size $\alpha(G)$, or the independence number is an important graph characteristics.
The method we use to compute this quantity is based on the following observations.
Let $x=\infty$, the independence polynomial in the previous section becomes
\begin{equation}
I(G, \infty) = \lim_{x\rightarrow \infty}\sum_{k=0}^{\alpha(G)}a_k x^k = a_{\alpha(G)} \infty^{\alpha(G)},
\end{equation}
where the lower-order terms vanish. We can thus replace the polynomial type $a = (a_0, a_1, \ldots, a_k)$ with a new type with two fields: the largest exponent $k$ and its coefficient $a_k$.
\purple{I was initially confused
because I thought 0 = (0, ?) where “?” could be anything positive since it's multiplied
by zero. It confused me that ? could equal $-\infty$ since I thought you defined your ring for
k positive, since a polynomial has positive exponents. If you have negative exponents I
thought it is a Laurent polynomial or something. After re-reading I see why this doesn't
matter, though.}\blue{I let $\mymathbb{0} = (1, -\infty)$ now, it is still correct, is it easier for understanding or I change it back?}
From this, we can define a new algebra as
\begin{equation}
    \eqname{P1}
\begin{split}
    a_x\infty^x \oplus a_y\infty^y &= \begin{cases}
        (a_x + a_y)\infty^{\max(x,y)}, & x = y\\
        a_y\infty^{\max(x,y)}, & x < y\\
        a_x\infty^{\max(x,y)}, & x > y
    \end{cases}, \\
    a_x\infty^x \odot a_y\infty^y &= a_x a_y\infty^{x+y}\\
    \mymathbb{0} &= 1\infty^{-\infty}\\
    \mymathbb{1} &= 1\infty^{0}.
\end{split}%& \text{\rlap{(P1)}}
\label{eq:countingtropical}
\end{equation}
To implement this algebra programmatically, we create a data type with two fields $(x, a_x)$ to store the MIS size and its counting,
and define the above operations and constants correspondingly.
If one is only interested in finding the MIS size, one can drop the counting field.
The algebra of the exponents becomes the max-plus tropical algebra~\cite{Maclagan2015, Moore2011}:
\begin{equation}\eqname{T}
    \begin{split}
        x \oplus y &= \max(x,y)\\
        x \odot y &= x + y\\
        \mymathbb{0} &= -\infty\\
        \mymathbb{1} &= 0.
    \end{split}\label{eq:tropical}
\end{equation}
Algebra \Eq{eq:tropical} and \Eq{eq:countingtropical} are the same as those used in Liu et al.~\cite{Liu2021} to calculate and count spin glass ground states.
For independent set calculations here, the vertex tensor and edge tensor becomes:
\begin{equation}
    W^{\rm T} = \left(\begin{matrix}
        \mymathbb{1} \\
        \infty
    \end{matrix}\right),   
    \qquad \qquad
        B^{\rm T} = \left(\begin{matrix}
        \mymathbb{1}  & \mymathbb{1} \\
        \mymathbb{1} & \mymathbb{0}
    \end{matrix}\right).
\end{equation}

@@

\begin{equation}
\eqname{SPT}
\begin{split}
    s \oplus t &= (\texttt{SUM}, /, s, t)\\
    s \odot t &= (\texttt{PROD}, /, s, t)\\
    \mymathbb{0} &= (\texttt{ZERO}, /, /, /)\\
    \mymathbb{1} &= (\texttt{LEAF}, 0^{\otimes |V|}, /, /).
\end{split}\label{eq:expr}
\end{equation}
\purple{It might be nice to have a picture of an example tree if you have time. However,
it is pretty clear from the definition, it would just help to visualize. Do you ever use the
zero element?}\blue{now we have an example, but it does not include a zero element, do you think I should use a more concrete one instead of an abstract one?}
The vertex tensor and edge tensor become
\begin{equation}
    W^{\rm SPT}((\texttt{LEAF},\boldsymbol{e}_i,/,/)) = \left(\begin{matrix}
        \mymathbb{1} \\
        (\texttt{LEAF},\boldsymbol{e}_i,/,/)
    \end{matrix}\right),   
    \qquad
        B^{\rm SPT} = \left(\begin{matrix}
        \mymathbb{1}  & \mymathbb{1} \\
        \mymathbb{1} & \mymathbb{0}
    \end{matrix}\right).
\end{equation}

This algebra is a commutative semiring because we define the equivalence of two sum-product trees by comparing their expanded (using \Eq{eq:set}) forms, rather than their storages.
Although one is probably unable to enumerate the configurations represented by this tree due to the extremely large configuration space size,
but he can directly sample configurations from this tree.
This is because once we have this sum-product representation of the configuration space,
either generate by using this algebra directly or serve as the coefficients of (truncated) polynomials,
one can compute the number of configurations in each branch rigorously and efficiently.
\purple{To clarify - is the sum-product tree now the coefficients
on the polynomial? I am a little confused on how we would, for instance, use this to
compute the MISs}\blue{added one more descriptive sentence, does it resolve your concern?}
To generate samples, one just descends this sum product tree to left or right sibling with probability decided by the size of each branch while doing the computation specified by the first field.
\xpurple{How do we compute the size of each
branch?}
\begin{example}
Let us consider the following sum product tree 
\begin{equation*}
    (\texttt{SUM}, /, (\texttt{PROD}, /, A, B), (\texttt{SUM}, /, C, D)),
\end{equation*}
where siblings $A, B, C$ and $D$ can be any four types of nodes.
It can be represented diagramatically as the following.

\centerline{\begin{tikzpicture}[]
    \def\dx{0};
    \def\a{1.0};
    \def\b{0.5};
    \def\ya{-1.0};
    \def\yb{-0.8};
    \node[] at (\dx, 0) (P1) {$\oplus$};
    \node[] at (\dx+\a, \ya) (P2) {$\oplus$};
    \node[] at (\dx-\a, \ya) (M1) {$\odot$};
    \node[] at (\dx-\a-\b, \ya+\yb) (A) {$A$};
    \node[] at (\dx-\a+\b, \ya+\yb) (B) {$B$};
    \node[] at (\dx+\a+\b, \ya+\yb) (C) {$C$};
    \node[] at (\dx+\a-\b, \ya+\yb) (D) {$D$};
    \draw[] (P1) -- (M1);
    \draw[] (P1) -- (P2);
    \draw[] (P2) -- (C);
    \draw[] (P2) -- (D);
    \draw[] (M1) -- (A);
    \draw[] (M1) -- (B);
\end{tikzpicture}}

The left and right sibling of root node has size $|A|\times |B|$ and $|C|+|D|$ respectively,
and since the root node has type \texttt{SUM}, its size can be computed as $|A|\times |B| + |C|+|D|$.
One can compute the sizes of $A, B, C$ and $D$ recursively until the program meets either a the \texttt{LEAF} node or a \texttt{ZERO} node, which has a known size of $1$ or $0$ respectively.
\end{example}

@@

Figure~\ref{fig:benchmark}(a) shows the time and space complexity of tensor network contraction for different graph sizes, where the space complexity is the same as the treewidth of the problem graph.
The contraction order is obtained using the local search algorithm in Ref.~\cite{Kalachev2021}.
In practice, slicing technique~\cite{Kalachev2021} is used for graphs with treewidth greater than 27 to fit the computation into a 16GB memory.
One can see that all the computing times in figure~\ref{fig:benchmark}(b), (c) and (d) have a strong correlation with the treewidth\purple{with the number
of vertices, and therefore the treewidth}.
\purple{Does this say anything about how
optimized the contraction order is? If so, it might be worth explicitly saying. Some of
the quantities in panel d look like they might be superexponential in the number of
vertices (tree-width).}
Among these benchmarks, computational tasks with data types \texttt{T (CPU)},
\texttt{$\mathbbm{R}$ (CPU)}, \texttt{$\mathbbm{R}$ (GPU)}, \texttt{$\mathbbm{C}$ (CPU)}, \texttt{$\mathbbm{C}$ (GPU)} and \texttt{T+bounding (CPU)} can utilize fast BLAS functions to evaluate tensor contractions, hence 
are much faster comparing to non-BLAS element types in the same category.
\purple{It might be worth explicitly saying
the data types the first time they're mentioned (for example, I don't know what T is.
Also, it might be worth adding a couple descriptor words on bounding.}
GPU computes much faster than CPU in all cases when the problem scale is large enough such that the actual computing time is comparable or larger than the launching overhead of CUDA kernels.
Most algebras can be computed on a GPU, except those requiring dynamic sized structures, i.e.\ types with algebra defined in \Eq{eq:polynomial}, \Eq{eq:set}, \Eq{eq:ext-tropical} and \Eq{eq:expr}.
In figure~\ref{fig:benchmark}(c), one can see the Fourier transformation based method is the fastest in computing the independence polynomial,
but it may suffer from round-off errors (\App{sec:fft})\purple{Can the round-off error thing be seen from Figure c?}\blue{it is hard, so I added a ref to the appendix, is it more informative?}. The finite field (GF$(p)$) approach is the only method that does not have round-off errors and can be run on a GPU.
In figure~\ref{fig:benchmark}(d), one can see the technique to bound the enumeration space improves the performance for more than one order of magnitude in enumerating the MISs.
Bounding can also reduce the memory usage significantly, without which the largest computable graph size is only $\sim150$ on a device with 32GB main memory.

@@

A theoretically interesting number that can be computed using the number of independent sets is the entropy constant for the hardcore lattice gases on these graphs, which can tell us thermal dynamic properties of these lattices at the high temperature limit.
\purple{Why is it useful and
what does it tell us about the physical system?}\blue{It is probably useless, so: useful -> theoretically interesting}
For the square grid graphs, it is called the \textit{hard square entropy constant} (\href{https://oeis.org/A085850}{OEIS A085850}), which is defined as $\lim_{L\rightarrow \infty} F(L,L)^{1/L^2}$, where $F(L,L)$ is the number of independent sets of a given lattice dimensions $L \times L$. This quantity arises in statistical mechanics of hard-square lattice gases~\cite{Baxter1980, Pearce1988} and is used to understand phase transitions for these systems. This entropy constant is not known to have an exact representation, but it is accurately known in many digits. Similarly, we can define entropy constants for other lattice gases. In \Fig{fig:hardsquare}, we look at how $F(L,L)^{1/\lfloor pL^2\rceil}$ scales as a function of the grid size $L$ for all types of graphs shown in Figure~\ref{fig:lattices}. Our results match the known results for the non-disordered square grid and King's graphs. For disordered square grid and King's graphs with a filling factor $p=0.8$, we randomly sample 1000 graph instances. To our knowledge, the entropy constants for these disordered graphs have not been studied before. Interestingly, the variations due to different random instances are negligible for this entropy quantity.

@@

\begin{figure}[th]
   \centering
\centerline{\begin{tikzpicture}[]
    \scriptsize
    \node[draw,fill=lime!80,fill opacity=1, text opacity=1.0,ellipse,minimum width=2cm, minimum height=1cm,inner sep=0pt] at (0, 2.5) (R) {$\mathbbm{R}$};
    \def\dx{-3};
    \node[draw,fill=teal!50,fill opacity=1, text opacity=1.0,ellipse,minimum width=5cm, minimum height=3cm,inner sep=0pt] at (\dx, 0) (PN) {\hspace{-3.5cm}Polynomial};
    \node[draw,fill=brown!75,fill opacity=1, text opacity=1.0,ellipse,minimum width=3.5cm, minimum height=1.0cm,inner sep=0pt] at (\dx, 0.7) (P1) {\hspace{-1.0cm}Largest order};
    \node[draw,fill=brown!40,fill opacity=1, text opacity=1.0,ellipse,minimum width=3.5cm, minimum height=1.0cm,inner sep=0pt] at (\dx, -0.4) (P2) {\hspace{-0.8cm}2nd largest order};
    \node[draw,fill=brown,fill opacity=1, text opacity=1.0,ellipse,minimum width=0.8cm, minimum height=0.3cm,inner sep=0pt] at (\dx, 1.0) (T) {};
    \node[draw,fill=brown!50,fill opacity=1, text opacity=1.0,ellipse,minimum width=0.8cm, minimum height=0.3cm,inner sep=0pt] at (\dx, -0.1) (T2) {};
    \node at (\dx, -1.2) {$\ldots$};
    \node[above = 1cm] at (T) (textT) {Tropical};
    \node[below = 2cm, left=4cm] (textT2) {Extended Tropical};
    \draw[black,-latex] (textT) -- (T);
    \draw[black,-latex] (textT2) -- (T2);
